### 2021-10-02 documents/sections/verbs.tex
added 'bathe' data, rendering tables

\subsection{\rc{(ət-)jəpɨ} \qu{to come}}
\label{sec:come}
\begin{table}
	\centering
	\caption{Reflexes of \qu{to come} \parencites[150]{alves2017arara}[265]{ikpengpacheco2001}[4]{meira2003bakairi}[294]{triomeira1999}[168]{meira1998proto}[178]{robayo2000avance}[37]{koehn1986apalai}[429]{courtz2008carib}[438]{maquiritaricaceres2011}[160]{stegeman2014akawaio}[125]{akawaiocaesar2003}[299, 415]{cruz2005fonologia}[102]{alvarez2000construcciones}[32]{macushiabbott1991}[65]{panarepayne2013}[68]{mendez1959yawarana}[74]{muller1975mapoyo}[182; p.c., Spike Gildea]{meira2005southern}}
    \label{tab:come}
\begin{tabular}[t]{@{}ll@{}}
\toprule
Language &                               Form \\
\midrule
\kaxui    &  \obj{johɨ}, \obj{ehɨ}, \obj{oohɨ} \\
\arara    &             \obj{ebɨ}, \obj{odebɨ} \\
\ikpeng   &                         \obj{arep} \\
\bakairi  &                         \obj{əewɨ} \\
\trio     &              \obj{epɨ}, \obj{əepɨ} \\
\akuriyo  &                         \obj{eepɨ} \\
\carijo   &                          \obj{ehɨ} \\
\apalai   &                         \obj{oepɨ} \\
\kalina   &                          \obj{opɨ} \\
\maqui    &                          \obj{ehə} \\
\akawaio  &            \obj{jepɨ}, \obj{əsipɨ} \\
\ingariko &               \obj{jepə}, \obj{jə} \\
\patamona &             \obj{jəpɨ}, \obj{jepɨ} \\
\pemon    &                         \obj{jepɨ} \\
\macushi  &                          \obj{ipɨ} \\
\panare   &                          \obj{əpɨ} \\
\yawarana &                          \obj{əpɨ} \\
\mapoyo   &                          \obj{epɨ} \\
\uxc      &                           \obj{ee} \\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{table}
	\centering
	\caption{Reflexes of \rc{ɨpɨtə} \qu{to go down}}
	\label{tab:descend_cog}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
Language & Verb & Class & Source \\
\mymidrule
\kaxui & \obj{ɨhɨto} & \gl{s_p_} & \perscomm{Spike Gildea} \\
\hixka & \obj{hto} & ? & \textcite[196]{hixkaryanaderby1979}\\
\waiwai & \obj{hto} & -- & \textcite[55]{waiwaihawkins1998}\\
\PPek & \rc{ɨptə} & \gl{s_a_} & \cref{sec:pekodian}\\
\PTir & \rc{ɨhtə}  (\gl{1} \rc{p-})& \gl{s_a_} & \cref{sec:taranoan} \\
\carijo & \obj{ehɨtə} & -- & \textcite[118]{guerrero2019carijo}\\
\wayana & \obj{ɨptə} & \gl{s_a_}/\gl{s_p_} & \textcite[44]{camargo2010wayana}\\
\apalai & \obj{ɨhto} & \gl{s_p_} & \textcite[99]{camargo2002lexico} \\
\kalina & (\obj{onɨʔto}) & \gl{s_a_} & \textcite[263]{courtz2008carib}\\
\maqui & \obj{əʔtə} & \gl{s_p_} & \textcite[450]{maquiritaricaceres2011}\\
\kapon & (\obj{uʔtə}) & -- & \textcite[139]{stegeman2014akawaio}\\
\pemon & (\obj{uʔtə}) & -- & \textcite[139]{alvarez2008clausulas}\\
\macushi & (\obj{autɨ}) & -- & \textcite[34]{macushiabbott1991}\\
\panare & \obj{əhtə} & \gl{s_a_} & \textcite[88]{mattei1994diccionario}\\
\yawarana & \obj{əhtə} & -- & \textcite[68]{mendez1959yawarana}\\
\waimiri & \obj{ɨtɨ} & -- & \textcite[58]{bruno1996dictionary}\\
	\mybottomrule
	\end{tabular}
\end{table}

\begin{table}
	\centering
	\caption{Reflexes of \rc{ɨpɨtə} \qu{to go down} \parencites[294, 325, 331]{triomeira1999}[196]{hixkaryanaderby1979}[55]{waiwaihawkins1998}[118]{guerrero2019carijo}[44]{camargo2010wayana}[99]{camargo2002lexico}[263]{courtz2008carib}[450]{maquiritaricaceres2011}[139]{stegeman2014akawaio}[139]{alvarez2008clausulas}[34]{macushiabbott1991}[88]{mattei1994diccionario}[68]{mendez1959yawarana}[58]{bruno1996dictionary}[84]{gildea1994akuriyo}[153]{alves2017arara}[137]{von1892bakairi}[116; p.c., Angela Fabíola Alves Chagas, David Felipe Guerrero, Spike Gildea]{meira1998proto}}
	\label{tab:XXX}
\begin{tabular}[t]{@{}lll@{}}
\toprule
Language &            Form &      Class \\
\midrule
\PPar     &     \obj{ɨɸɨto} &  \gl{s_p_} \\
\kaxui    &     \obj{ɨhɨto} &  \gl{s_p_} \\
\hixka    &       \obj{hto} &          ? \\
\waiwai   &       \obj{hto} &          – \\
\PPek     &      \obj{ɨptə} &  \gl{s_a_} \\
\arara    &     \obj{iptoŋ} &  \gl{s_a_} \\
\ikpeng   &     \obj{iptoŋ} &          ? \\
\bakairi  &     \obj{ɨtəgɨ} &  \gl{s_a_} \\
\PTir     &      \obj{ɨhtə} &  \gl{s_a_} \\
\trio     &      \obj{ɨhtə} &  \gl{s_a_} \\
\akuriyo  &       \obj{ɨtə} &  \gl{s_a_} \\
\carijo   &     \obj{ehɨtə} &          – \\
\wayana   &      \obj{ɨptə} &  \gl{s_p_} \\
\apalai   &      \obj{ɨhto} &  \gl{s_p_} \\
\kalina   &  (\obj{onɨʔto}) &  \gl{s_a_} \\
\maqui    &      \obj{əʔtə} &  \gl{s_p_} \\
\kapon    &    (\obj{uʔtə}) &          – \\
\pemon    &    (\obj{uʔtə}) &          – \\
\macushi  &    (\obj{autɨ}) &          – \\
\panare   &      \obj{əhtə} &  \gl{s_a_} \\
\yawarana &      \obj{əhtə} &          – \\
\waimiri  &       \obj{ɨtɨ} &          – \\
\bottomrule
\end{tabular}
\end{table}

A comparative view \pcref{tab:descend_cog} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb.
It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.

@@

\begin{table}
	\centering
	\caption{Intransitive and transitive \qu{to bathe}}
    \label{tab:bathe}
	\begin{tabular}{@{}llll@{}}
	\mytoprule
	Language & \gl{intr} & \gl{tr} & Source\\
	\mymidrule
\kaxui & \obj{eehɨ} & \obj{ɨhɨ} & \perscomm{Spike Gildea}\\
\hixka & \obj{ewehɨ} & \obj{ɨhɨ} & \textcites[198]{hixkaryanaderby1979}\\
\waiwai & \obj{ejepu} & \obj{pɨ} & \textcites[166, 192]{waiwaihawkins1998}\\
\arara & \obj{ibɨ} & \obj{ɨp} & \textcites[150, 162]{alves2017arara}\\
\ikpeng & \obj{ip} & \obj{ɨp} & \textcites[103]{ikpengpacheco1997}[123]{campetela1997analise}\\
\bakairi & \obj{i} & \obj{ɨ} & \textcites[4]{meira2003bakairi}[285]{meira2005bakairi}\\
\trio & \obj{epɨ} & \obj{pɨ} & \textcites[697]{triomeira1999}\\
\akuriyo & \obj{epɨ} & \obj{pɨ} & \textcites[87]{gildea1994akuriyo}\\
\wayana & \obj{epɨ} & \obj{upɨ} & \textcites[24, 52]{camargo2010wayana}\\
\apalai & \obj{epɨ} & \obj{pɨ} & \textcites[218]{meira2000split}\\
\kalina & \obj{ekupi} & \obj{kupi} & \textcites[304]{courtz2008carib}\\
\maqui & \obj{eʔhi} & \obj{ɨhɨ} & \textcites[439, 454]{maquiritaricaceres2011}\\
\kapon & \obj{ekuʔpi} & \obj{kuʔpi} & \textcites[37]{stegeman2014akawaio}\\
\pemon & \obj{ekupɨ} & \obj{pɨ} & \textcites[34, 129]{pemondearmellada1944dic}\\
\panare & \obj{akupɨ} & \obj{ɨpɨ}/\obj{kupɨ} & \textcites[8, 294]{mattei1994diccionario}\\
\mybottomrule
	\end{tabular}
\end{table}

@@

>>>>

\subsection{\rc{(ət-)jəpɨ} \qu{to come}}
\label{sec:come}

\input{floats/come}

@@

\input{floats/godown}

A comparative view \pcref{tab:godown} shows that while a form \rc{ɨpɨtə} can be reconstructed to \PC, the class of this verb is unclear at first sight.
The reflexes are fairly evenly split between \gl{s_a_} and \gl{s_p_} -- in those languages that preserve the split-\gl{s} system.
In one language, \wayana, the verb shows traits of both classes, leading us to consider it a \dbqu{mixed} verb.
It takes the first and second person \gl{s_p_} markers \obj{j-} and \obj{əw-} \parencite[200]{wayanatavares2005}, but the \gl{1+2}\gl{s_a_} marker \obj{kut-} \parencite[206]{wayanatavares2005}.
It also shows the \gl{s_a_} class marker \obj{w-} in nominalizations \exref{waygodown.way-73}, but behaves like an \gl{s_p_} verb in taking a second person prefix in imperatives \exref{waygodown.way-71}.

@@

\input{floats/bathe}

@@

\input{floats/defecate}
